\section{Related Work}
\label{sec:related}

BiBoP allocators.

\paragraph{Randomized memory management.} Several previous memory managers have employed randomization,
primarily for fault tolerance and
security~\cite{Novark:2010:DSH:1866307.1866371, 1134000, 1346296,
  1250736}. DieHard uses randomized memory allocation to provide
\emph{probabilistic memory safety}, which enables (probabilistically)
correct execution in the face of memory usage errors like heap buffer
overflows and use-after-free~\cite{1134000}. Archipelago extends
DieHard by placing each object in a randomly-selected page in a 64-bit
address space, significantly extending its ability to tolerate memory
errors~\cite{1346296}. Archipelago reduces its footprint by copying
unused objects into a conventional heap and discards the backing
physical page; it then uses virtual memory page protection to
intercept reads and re-instantiates the physical page on
demand. DieHarder extends DieHard to provide security against attacks
that exploit memory errors~\cite{Novark:2010:DSH:1866307.1866371},
while Exterminator leverages a randomized memory allocator to perform
statistical inference and automatically fix the application by
generating ``runtime patches''~\cite{1250736}. Stabilizer randomizes
the placement of code, stack frames, and heap objects to enable
statistically sound performance evaluation~\cite{stabilizer:asplos13};
we adopt Stabilizer's shuffling heap approach. Unlike all of these,
meshing uses randomization to enable it to effectively compact objects.

\paragraph{Virtual memory approaches.}
Appel and Li outline a number of approaches for using virtual memory
in user programs~\cite{Appel:1991:VMP:106972.106984}. Numerous memory
managers have exploited virtual memory primitives for a variety of
purposes~\cite{Novark:2010:DSH:1866307.1866371,1346296,Appel:1988:RCC:53990.53992,Boehm:1991:MPG:113445.113459}. The
only memory manager that merges virtual pages is Hound, a memory leak
and ``bloat'' detector~\cite{1542521}. Hound uses periodic page
protection to track accesses and thus identify stale objects. By
allocating objects sequentially so they are age-segregated, it
eventually isolates leaked objects and bloat on their own pages. To
avoid mixing young and old objects, Hound does not reuse freed slots
on a page until it becomes completely empty. Because this lack of
reuse could lead to catastrophic memory consumption, Hound merges
virtual pages onto physical pages as a backstop. Meshing adapts this
approach and combines it with a randomized algorithm that lets it
effectively merge pages regardless of allocation pattern. Unlike
meshing, Hound's deterministic allocation strategy cannot guarantee
successful compaction and is best-effort only.
